Wigglyhedra

### 2024-05-31 draft/wigglyhedra.tex
Moved categorical section; started writing some propositions to prove.

\subsection{Curves in a disk and a triangulated category associated to a quiver}
Curves up to isotopy in a disk with marked points constitute a geometric model of certain objects in a triangulated category.
This correspondence, due to Khovanov--Seidel~\cite{kho.sei:02}, and will enable us to build categorical interpretations of the combinatorial results obtained in this paper.
The aim of this section is to give a concise account of this correspondence and some of its useful features; we refer the reader to~\cite{kho.sei:02} for details.

The category of interest is a subcategory of \(D^b(A_m\text{-mod})\) for a certain finite-dimensional algebra \(A_m\).
That is, a subcategory of the bounded derived category of (finitely generated) \(A_m\)-modules.
The algebra \(A_m\), called the \emph{zigzag algebra}, is defined as a quotient of the path algebra of the quiver \(Q_m\) with \(m\) vertices depicted in~\cref{fig:am-quiver}.
\begin{figure}[h]
  \centering
  \begin{tikzcd}
    1 \arrow[bend left]{r} & 2 \arrow[bend left]{r}\arrow[bend left]{l}& \cdots \arrow[bend left]{r}\arrow[bend left]{l}& m \arrow[bend left]{l}
  \end{tikzcd}
  \caption{The quiver \(Q_m\). The algebra \(A_m\) is a quotient of the path algebra \(kQ_m\).}
  \label{fig:am-quiver}
\end{figure}
Let \(kQ_m\) be the path algebra of \(Q_m\), graded by path length.
Let \(e_i\) denote the idempotent path at the vertex \(i\), and let \((i|j)\) denote the arrow from vertex \(i\) to a (neighbouring) vertex \(j\).
A longer path, which is a composition of arrows, will be denoted by \((i_1|i_2|\cdots|i_n)\).

The zigzag algebra \(A_m\) is the quotient of \(kQ_m\) by the following relations:
\begin{itemize}
\item set \((i|j|l) = 0\) whenever \(i \neq l\);
\item set \((i|j|i) = (i|l|i)\) for every \(j\) and \(l\) such that there are arrows between \(i\) and \(j\) and between \(i\) and \(l\).
\end{itemize}
See~\cite[Section 3]{hue.kho:01} for further details, and~\cite[Remark 6.6]{bap.deo.lic:20} for a more detailed discussion.
Let \(P_i\) denote the indecomposable projective \(A_m\)-module defined as \(P_i = A_m\cdot e_i\).
For any integer \(j\), let \(P_i \langle j \rangle\) denote its graded shift by \(j\).
Set \(\mathcal{C}_m\) to be the smallest triangulated subcategory of \(D^b(A_m\text{-mod})\) containing all the objects \(P_i\langle j \rangle\) (see, e.g.,~\cite[Section 2.3.3]{bap.deo.lic:22} for further discussion).

\subsection{The Khovanov--Seidel curve model}
Now consider \(n+2\) marked points on a disk, arranged on a straight horizontal line, and labelled \(0,1, \ldots, n+1\).
We consider isotopy classes of curves whose endpoints lie on distinct marked points, and which do not cross themselves or any marked point in their interior.
For convenience, we also refer to such an isotopy class as a \emph{curve}.
The Khovanov--Seidel correspondence (explained in Section 4 of~\cite{kho.sei:02}) associates objects of \(\mathcal{C}_{n+1}\) (up to grading\asilata{Should this say homological?} shift) to curves on a configuration of \(n+2\) marked points arranged in a straight line.

@@

\begin{definition}
Two wiggly arcs~$(i,j,A,B)$ and~$(i',j',A',B')$ are 
\begin{itemize}
\item \defn{non pointed} if~$i = j'$ or~$i' = j$,
\item \defn{crossing} if~$(A \cap B') \cup (\{i,j\} \cap B') \cup (A \cap \{i',j'\}) \ne \varnothing \ne (A' \cap B) \cup (\{i',j'\} \cap B) \cup (A' \cap \{i,j\})$, that is, if the corresponding curves cross,
\item \defn{incompatible} if they are non pointed or crossing, and \defn{compatible} otherwise.
\end{itemize}
%
\begin{figure}[h]
\centerline{\includegraphics[scale=1.3]{incompatible}}
\caption{Some incompatible wiggly arcs: non pointed (left) and crossing (right).}
\label{fig:incompatible}
\end{figure}
\end{definition}

@@

>>>>

@@

\begin{definition}\label{def:compatible}
Two wiggly arcs~$(i,j,A,B)$ and~$(i',j',A',B')$ are 
\begin{itemize}
\item \defn{non pointed} if~$i = j'$ or~$i' = j$,
\item \defn{crossing} if~$(A \cap B') \cup (\{i,j\} \cap B') \cup (A \cap \{i',j'\}) \ne \varnothing \ne (A' \cap B) \cup (\{i',j'\} \cap B) \cup (A' \cap \{i,j\})$, that is, if the corresponding curves cross,
\item \defn{incompatible} if they are non pointed or crossing, and \defn{compatible} otherwise.
\end{itemize}
%
\begin{figure}[h]
\centerline{\includegraphics[scale=1.3]{incompatible}}
\caption{Some incompatible wiggly arcs: non pointed (left) and crossing (right).}
\label{fig:incompatible}
\end{figure}
\end{definition}

@@

\section{Categorical interpretation of wiggly pseudotriangulations}
The aim of this section is to explain a categorical interpretation of wiggly arcs.
We recall results of Khovanov--Seidel~\cite{kho.sei:02} that realise wiggly arcs as objects of a triangulated category, such that intersections between wiggly arcs correspond to morphism in this category.
We also discuss categorical conditions for two wiggly arcs to be compatible as in~\cref{def:compatible}.

% Curves up to isotopy in a disk with marked points constitute a geometric model of certain objects in a triangulated category.
% This correspondence, due to Khovanov--Seidel~\cite{kho.sei:02}, and will enable us to build categorical interpretations of the combinatorial results obtained in this paper.
% The aim of this section is to give a concise account of this correspondence and some of its useful features; we refer the reader to~\cite{kho.sei:02} for details.

\subsection{A triangulated category associated to a quiver}
We begin by describing the category of interest, which is a subcategory of \(D^b(A_m\text{-mod})\) for a certain finite-dimensional algebra \(A_m\).
That is, a subcategory of the bounded derived category of (finitely generated) \(A_m\)-modules.
The algebra \(A_m\), called the \emph{zigzag algebra}, is defined as a quotient of the path algebra of the quiver \(Q_m\) with \(m\) vertices depicted in~\cref{fig:am-quiver}.
\begin{figure}[h]
  \centering
  \begin{tikzcd}
    1 \arrow[bend left]{r} & 2 \arrow[bend left]{r}\arrow[bend left]{l}& \cdots \arrow[bend left]{r}\arrow[bend left]{l}& m \arrow[bend left]{l}
  \end{tikzcd}
  \caption{The quiver \(Q_m\). The algebra \(A_m\) is a quotient of the path algebra \(kQ_m\).}
  \label{fig:am-quiver}
\end{figure}
Let \(kQ_m\) be the path algebra of \(Q_m\), graded by path length.
Let \(e_i\) denote the idempotent path at the vertex \(i\), and let \((i|j)\) denote the arrow from vertex \(i\) to a (neighbouring) vertex \(j\).
A longer path, which is a composition of arrows, will be denoted by \((i_1|i_2|\cdots|i_n)\).

\begin{definition}
  The zigzag algebra \(A_m\) corresponding to the quiver \(Q_m\) depicted in~\cref{fig:am-quiver} is the quotient of \(kQ_m\) by the following relations:
  \begin{itemize}
  \item \((i|j|l) = 0\) whenever \(i \neq l\);
  \item \((i|j|i) = (i|l|i)\) for every \(j\) and \(l\) such that there are arrows between \(i\) and \(j\) and between \(i\) and \(l\).
  \end{itemize}
\end{definition}

See~\cite[Section 3]{hue.kho:01} for further details on the definition, and~\cite[Remark 6.6]{bap.deo.lic:20} for more discussion.
Let \(P_i\) denote the indecomposable projective \(A_m\)-module defined as \(P_i = A_m\cdot e_i\).
For any integer \(j\), let \(P_i \langle j \rangle\) denote its graded shift by \(j\).
Set \(\mathcal{C}_m\) to be the smallest triangulated subcategory of \(D^b(A_m\text{-mod})\) containing all the objects \(P_i\langle j \rangle\) (see, e.g.,~\cite[Section 2.3.3]{bap.deo.lic:22} for further discussion on the construction).

\subsection{The Khovanov--Seidel curve model}
In this section we briefly recall the procedure to recover an object of \(\mathcal{C}_{n+1}\) from a wiggly arc on \(n+2\) points on a line.
In fact, the procedure works on more general arcs as well --- namely, those that are not necessarily \(x\)-monotone.
All the results of this section are due to Khovanov and Seidel~\cite{kho.sei:02}.

Consider \(n+2\) points arranged in a straight horizontal line, and labelled \(0,1, \ldots, n+1\) as in~\cref{sec:combinatorics}.
In addition to the wiggly arcs, consider isotopy classes of curves whose endpoints lie on distinct marked points, and which do not cross themselves or any marked point in their interior.
For convenience, we refer to such an isotopy class as an \emph{arc}.

The Khovanov--Seidel correspondence (explained in Section 4 of~\cite{kho.sei:02}) associates objects of \(\mathcal{C}_{n+1}\) (up to grading\asilata{Should this say homological?} shift) to curves on a configuration of \(n+2\) marked points arranged in a straight line.

\subsection{Categorical conditions on compatibility of arcs}

\begin{proposition}
  Let \(X \in \mathcal{C}_{n+1}\) be a spherical object.
  Consider its cohomology filtration
  \[
    \begin{tikzcd}[column sep=0.3cm]
      0 = X_0 \arrow{rr} && X_1 \arrow{dl} \arrow{rr} && \cdots \arrow{rr} \arrow{dl} && X_{k-1} \arrow{rr} \arrow{dl} && X_k = X \arrow{dl}\\
      & H^1(X) \arrow[dashed]{ul}{+1} && H^2(X) \arrow[dashed]{ul}{+1} && \cdots \arrow[dashed]{ul} &&H^k(X) \arrow[dashed]{ul}{+1}
    \end{tikzcd}
  \]
  Consider the set of objects \(\{H^j(X)\}\) taken without multiplicity, and let \(T\) be corresponding set of wiggly curves.
  We have the following results.
  \begin{enumerate}
  \item 
    \(T\) is a subset of a wiggly pseudotriangulation.
  \item \(T\) contains at most one of the two irrelevant arcs.
  \end{enumerate}
\end{proposition}
